---
title: "Data Value Methodes Decisionnelles - TP6A "
output: html_notebook
---

TP6A Réseaux de neurones avec R. 
Auteur: Xavier VAN AUSLOOS
Version: 12/12/2020

## Create Test Data

```{r}
set.seed(42)
x<-runif(200, -10, 10)
y<-sin(x)
plot(x,y)
```
## Randomize the initial weights
A network function is made of three components: the network of neurons, the weight of each connection between neuron and the activation function of each neuron. For this example, we’ll use a feed-forward neural network and the logistic activation which are the defaults for the package nnet. We take one number as input of our neural network and we want one number as the output so the size of the input and output layer are both of one. For the hidden layer, we’ll start with three neurons. It’s good practice to randomize the initial weights, so create a vector of 10 random values, picked in the interval [-1,1].

```{r}
weight<-runif(10, -1, 1)
```

## Create Training Set and Test Set
Neural networks have a strong tendency of overfitting your data, meaning they become really good at describing the relationship between the values in your data set, but are not effective with data that wasn’t used to train your model. As a consequence, we need to cross-validate our model. Set the seed to 42, then create a training set containing 75% of the values in your initial data set and a test set containing the rest of your data.

```{r}
set.seed(42)
index<-sample(1:length(x),round(0.75*length(x)),replace=FALSE)
reg.train<-data.frame(X=x[index],Y=y[index])
reg.test<-data.frame(X=x[-index],Y=y[-index])
head(reg.train)
```

```{r}
head(reg.test)
```

## Usint nnet to create the model

Load the nnet package and use the function of the same name to create your model. Pass your weights via the Wts argument and set the maxit argument to 50. We want to fit a function which can have for output multiple possible values. To do so, set the linout argument to true. Finally, take the time to look at the structure of your model.

```{r}
library(nnet)
set.seed(42)
reg.model.1<-nnet(reg.train$X,reg.train$Y,size=3,maxit=50,Wts=weight,linout=TRUE)
```
Have a look at the model 
reg.model.1 ==> modellogical.  If true, the model frame is saved as componentmodelof the returnedobject.
```{r}
str(reg.model.1)
```

## Predict with RMSE
Predict the output for the test set and compute the RMSE of your predictions. Plot the function sin(x) and then plot your predictions.

```{r}
predict.model.1<-predict(reg.model.1,data.frame(X=reg.test$X))
str(predict.model.1)
```
RMSE: root mean square error (have a look here for details: https://www.kaggle.com/c/ashrae-energy-prediction/discussion/113064)

```{r}
rmse.reg<-sqrt(sum((reg.test$Y-predict.model.1)^2))
rmse.reg
```

```{r}
plot(sin, -10, 10)
points(reg.test$X,predict.model.1)
```
## The number of neurons in the hidden layer

The number of neurons in the hidden layer, as well as the number of hidden layer used, has a great influence on the effectiveness of your model. Repeat the exercises three to five, but this time use a hidden layer with seven neurons and initiate randomly 22 weights.

```{r}
set.seed(42)
reg.model.2<-nnet(reg.train$X,reg.train$Y,size=7,maxit=50,Wts=runif(22, -1, 1),linout=TRUE)

```

Have a look to model str(reg.model.2)

```{r}
str(reg.model.2)

```
```{r}
predict.model.2<-predict(reg.model.2,data.frame(X=reg.test$X))
str(predict.model.2)
```

```{r}
rmse.reg<-sqrt(sum((reg.test$Y-predict.model.2)^2))
rmse.reg
```
```{r}
plot(sin, -10, 10)
points(reg.test$X,predict.model.2)
```

# Using Neural Network for classification problem - Iris Dataset

Now let us use neural networks to solve a classification problem, so let’s load the iris data set! It is good practice to normalize your input data to uniformize the behavior of your model over different range of value and have a faster training. Normalize each factor so that they have a mean of zero and a standard deviation of 1, then create your train and test set.

```{r}
data<-iris

scale.data<-data.frame(lapply(data[,1:4], function(x) scale(x)))
scale.data$Species<-data$Species

index<-sample(1:nrow(scale.data),round(0.75*nrow(scale.data)),replace=FALSE)
clust.train<-scale.data[index,]
clust.test<-scale.data[-index,]
```

## Create Model on Iris Classifier

Use the nnet() and use a hidden layer of ten neurons to create your model. We want to fit a function which have a finite amount of value as output. To do so, set the linout argument to true. Look at the structure of your model. With classification problem, the output is usually a factor that is coded as multiple dummy variables, instead of a single numeric value. As a consequence, the output layer have as one less neuron than the number of levels of the output factor.

```{r}
set.seed(42)
clust.model<-nnet(Species~.,size=10,Wts=runif(83, -1, 1),data=clust.train)

```

##  Predict the value of the test set
Make prediction with the values of the test set.

```{r}
predict.model.clust<-predict(clust.model,clust.test[,1:4],type="class")
```

## Create teh confusion table
Create the confusion table of your prediction and compute the accuracy of the model.

```{r}
Table<-table(clust.test$Species ,predict.model.clust)
Table
```

```{r}
accuracy<-sum(diag(Table))/sum(Table)
accuracy
```

